## Nombre de turbines fissurées {#sec-glm-turbines_fissurees}

```{r}
#| echo: false

x <- c(400, seq(from = 1000, to = 4600, by = 400))
m <- c(39, 53, 33, 73, 30, 39, 42, 13, 34, 40, 36)
y <- c(2, 4, 3, 7, 5, 9, 9, 6, 22, 21, 21)
```

On considère les données suivantes, qui contiennent le nombre $Y_i$ de turbines sur $m_i$ qui ont été fissurées après $x_i$ heures d'opération.

```{r}
#| echo: false

tab <- data.frame(
  x_i = c(400, 1000, 1400, 1800, 2200, 2600, 3000, 3400, 3800, 4200, 4600),
  m_i = c(39, 53, 33, 73, 30, 39, 42, 13, 34, 40, 36),
  Y_i = c(2, 4, 3, 7, 5, 9, 9, 6, 22, 21, 21)
)

kable(
  tab,
  col.names = c("$x_i$", "$m_i$", "$Y_i$"),
  align = c("r", "r", "r"),
  escape = FALSE,
  booktabs = TRUE
)

```

**a)** En utilisant un GLM binomial avec lien canonique, dériver les estimateurs des paramètres lorsque $x_i$ est traité comme une variable exogène dichotomique avec 11 niveaux, et lorsque le prédicteur linéaire pour la donnée $i$ est $$\eta_i=\beta_0+\beta_i, \text{ pour } i=1,...,11,$$ avec la contrainte d'identifiabilité que $\beta_1=0$.

**b)** En utilisant \textsf{R} et un GLM binomial avec lien canonique, ajuster le modèle où le prédicteur linéaire est $$\eta_i=\beta_0+\beta_1 x_i, \text{ pour } i=1,...,11.$$ Donner les estimations des paramètres et leur écart-type.

**c)** Refaire (b) en utilisant un lien probit. Donner les estimations des paramètres et leur écart-type.

**d)** Refaire (b) en utilisant un lien log-log complémentaire. Donner les estimations des paramètres et leur écart-type.

**e)** Comparer les prévisions (et leurs mesures d'incertitude) sous les trois modèles ajustés en (b), (c) et (d) pour une turbine qui était en opération pour 2000 heures.

**f)** Tracer un graphique pour montrer si les modèles en (b), (c) et (d) ajustent bien (ou non) les données. Commenter.

::: {.callout-tip collapse="true" title="Solution complète"}
**a)** If $x_{i}$ is treated as a factor predictor with 11 levels, the linear predictor is written as 
$$
\eta_{i}=\beta_{0}+\beta_{i} \text{,  } i=1,...,11
$$ 

and $\beta_{1}=0$. The binomial density is the following : 

$$
f_{Y}(y_{i})=\begin{pmatrix} m_{i} \\ y_{i} \end{pmatrix}\pi_{i}^{y_{i}}(1-\pi_{i})^{m_{i}-y_{i}},
$$

which can be rewritten in a exponential family representation as : 
$$
f_{Y}(y_{i})=\exp\left[y_{i}\ln\left(\frac{\pi_{i}}{1-\pi_{i}}\right)+m\ln(1-\pi_{i})+\ln\begin{pmatrix} m_{i} \\ y_{i} \end{pmatrix}\right].
$$ 

Hence, the canonical parameter is $\theta_{i}=\ln\left(\frac{\pi_{i}}{1-\pi_{i}}\right)$ and the canonical link is the logit link. Thus, 

$$
\begin{aligned} \eta_{i}&=\ln\left(\frac{\pi_{i}}{1-\pi_{i}}\right)=\beta_{0}+\beta_{i} \\ \pi_{i}&=\frac{e^{\eta_{i}}}{1+e^{\eta_{i}}}= \frac{e^{\beta_{0}+\beta_{i}}}{1+e^{\beta_{0}+\beta_{i}}}\\ \end{aligned}
$$ The expression of the density in the reparametrization is then 

$$
\begin{aligned} f_{Y}(y_{i})&=\begin{pmatrix} m_{i} \\ y_{i} \end{pmatrix}\left(\frac{e^{\beta_{0}+\beta_{i}}}{1+e^{\beta_{0}+\beta_{i}}}\right)^{y_{i}}\left(\frac{1}{1+e^{\beta_{0}+\beta_{i}}}\right)^{m_{i}-y_{i}}\\ &=\begin{pmatrix} m_{i} \\ y_{i} \end{pmatrix}\frac{e^{y_{i}(\beta_{0}+\beta_{i})}}{(1+e^{\beta_{0}+\beta_{i}})^{m_{i}}}\\ \end{aligned}
$$ The likelihood $L$ and the log-likelihood $l$ are shown below : 

$$
\begin{aligned} L(\beta_{0},...,\beta_{11};y_{1},...,y_{11})&=\prod_{i=1}^{11} \begin{pmatrix} m_{i} \\ y_{i} \end{pmatrix}\frac{e^{y_{i}(\beta_{0}+\beta_{i})}}{(1+e^{\beta_{0}+\beta_{i}})^{m_{i}}}\\ \ell(\beta_{0},...,\beta_{11};y_{1},...,y_{11})&=\sum_{i=1}^{11} \left[\ln\begin{pmatrix} m_{i} \\ y_{i} \end{pmatrix}+y_{i}(\beta_{0}+\beta_{i})-m_{i}\ln(1+e^{\beta_{0}+\beta_{i}})\right]\\ \end{aligned}
$$ 

We have 

$$
\begin{aligned} \frac{\partial \ell}{\partial \beta_{0}}&=\sum_{i=1}^{11} \left[y_{i}-m_{i}\frac{e^{\beta_{0}+\beta_{i}}}{1+e^{\beta_{0}+\beta_{i}}}\right]\\ \frac{\partial \ell}{\partial \beta_{i}}&=y_{i}-m_{i}\frac{e^{\beta_{0}+\beta_{i}}}{1+e^{\beta_{0}+\beta_{i}}} \text{,  } i=2,...,11,\\ \end{aligned}
$$ and $\beta_{1}=0$ by constraint of the model. The maximum likelihood estimators for the parameters are derived by solving the system of equations $\frac{\partial \ell}{\partial \beta_{i}}=0,$ $i=0,...,11$ : 

$$
\begin{aligned} \sum_{i=1}^{11} \left[y_{i}-m_{i}\frac{e^{\hat{\beta}_{0}+\hat{\beta}_{i}}}{1+e^{\hat{\beta}_{0}+\hat{\beta}_{i}}}\right]&=0\\ y_{i}-m_{i}\frac{e^{\hat{\beta}_{0}+\hat{\beta}_{i}}}{1+e^{\hat{\beta}_{0}+\hat{\beta}_{i}}}&=0 \text{,  } i=2,...,11,\\ \Rightarrow \hat{\beta}_{0}+\hat{\beta}_{i}=\ln\left(\frac{y_{i}}{m_{i}-y_{i}}\right)& \text{,  } i=2,...,11,\\ \end{aligned}
$$  

Using the first equation and replacing $\hat{\beta}_{0}+\hat{\beta}_{i}$ by $\ln\left(\frac{y_{i}}{m_{i}-y_{i}}\right)$, 
$$
\begin{aligned} &y_{1}-m_{1}\frac{e^{\hat{\beta}_{0}}}{1+e^{\hat{\beta}_{0}}}+\sum_{i=2}^{11} \left[y_{i}-m_{i}\frac{\left(\frac{y_{i}}{m_{i}-y_{i}}\right)}{1+\left(\frac{y_{i}}{m_{i}-y_{i}}\right)}\right]=0\\ &y_{1}-m_{1}\frac{e^{\hat{\beta}_{0}}}{1+e^{\hat{\beta}_{0}}}+\sum_{i=2}^{11} \left[y_{i}-m_{i}\frac{y_{i}}{m_{i}}\right]=0\\ &y_{1}-m_{1}\frac{e^{\hat{\beta}_{0}}}{1+e^{\hat{\beta}_{0}}}=0\\ &\hat{\beta}_{0}=\ln\left(\frac{y_{1}}{m_{1}-y_{1}}\right)\\ &\hat{\beta}_{i}=\ln\left(\frac{y_{i}}{m_{i}-y_{i}}\right)-\hat{\beta}_{0}=\ln\left(\frac{y_{i}/(m_{i}-y_{i})}{y_{1}/(m_{1}-y_{1})}\right)  \text{,  } i=2,...,11.\\ \end{aligned}
$$ 

The estimates of the model parameters are easily found in R as follows : 
```{r} 
(beta0 <- log(y[1]/(m[1]-y[1]))) 
(beta <- c(0,log(y[-1]/(m[-1]-y[-1]))-beta0)) 
``` 

Hence, here 

$$
\hat{\beta}=(-2.9178, 0, 0.4122, 0.6152, 0.6740, 1.3083, 1.7138, 1.6185, 2.7636, 3.5239, 3.0179, 3.2542)^\top.
$$ 

As a consistency check following from the invariance property of maximum likelihood estimation, we can verify that the estimates of $\pi_{i}$ using the expit function are equal to the MLE estimates $\hat{\pi}_{i}=\frac{y_{i}}{m_{i}}$ : 
```{r} 
(pi <- exp(beta0+beta)/(1+exp(beta0+beta))) 
y/m 
```

**b)** The Binomial GLM model with logit link and the linear predictor $\eta_{i}=\beta_{0}+\beta_{1}x_{i}$, $i=1,...,11$ is fitted to the data using R and the command : 

```{r} 
#| eval: false 

glm(cbind(y,m-y) ~ x,family=binomial) 
``` 

The estimates of the parameters are $\hat{\beta}_{0}=-3.6070615$ and $\hat{\beta}_{1}=0.0009121$, with standard error $SE(\hat{\beta}_{0})=0.3533875$ and $SE(\hat{\beta}_{1})=0.0001084$.

**c)** The Binomial GLM model with probit link and the linear predictor $\eta_{i}=\beta_{0}+\beta_{1}x_{i}$, $i=1,...,11$ is fitted to the data using R and the command : 

```{r} 
#| eval: false 

glm(cbind(y,m-y) ~ x,family=binomial(link=probit)) 
``` 

The estimates of the parameters are $\hat{\beta}_{0}=-2.080$ and $\hat{\beta}_{1}=5.230\times10^{-4}$, with standard error $SE(\hat{\beta}_{0})=0.1852$ and $SE(\hat{\beta}_{1})=5.973\times10^{-5}$.

**d)** The Binomial GLM model with complementary log-log link and the linear predictor $\eta_{i}=\beta_{0}+\beta_{1}x_{i}$, $i=1,...,11$ is fitted to the data using R and the command: 

```{r} 
#| eval: false 

glm(cbind(y,m-y) ~ x,family=binomial(link=cloglog)) 
``` 

The estimates of the parameters are $\hat{\beta}_{0}=-3.360$ and $\hat{\beta}_{1}=7.480\times10^{-4}$, with standard error $SE(\hat{\beta}_{0})=0.3061$ and $SE(\hat{\beta}_{1})=8.622\times10^{-5}$.

**e)** Predictions can be found using the inverse of the link function. For the model with canonical link (model from b), we find that 
$$
\hat{y}_{2000}=\frac{e^{\hat{\beta}_{0}+2000\hat{\beta}_{1}}}{1+e^{\hat{\beta}_{0}+2000\hat{\beta}_{1}}}=0.1439472.
$$ 

Alternatively, the command `predict(modelb,data.frame(x=2000),type="response",se.fit=TRUE)` can be used to calculate the predictions and associated standard errors. The resulting predictions and standard errors are presented in the following table. The probability of developing fissures after 2000 hours of operations is 14.39 % according to model b, and the standard deviation is 2.08 %. This prediciton is quite comparable with the complementary log-log model (d), for which the estimated probability of developing fissures is 14.36 %, with a standard error of 2.03 %. The precision is slightly better in this model than the two others due to a smaller variance. The estimated probability with Model c, using the probit link, is higher at 15.06 %, with standard error of 2.06 %.

```{r}
#| echo: false


tab <- data.frame(
  Mesure = c("$\\hat{y}_{2000}$", "$SE(\\hat{y}_{2000})$"),
  "Model b (logit link)" = c(0.1439472, 0.02080256),
  "Model c (probit link)" = c(0.1506150, 0.02062154),
  "Model d (compl. log-log link)" = c(0.1436447, 0.02030803),
  check.names = FALSE
)

kable(
  tab,
  caption = "Predictions and Standard Errors for Predictions for the 3 Models",
  booktabs = TRUE,
  escape = FALSE,
  align = c("l", "c", "c", "c")
)

```

**f)** Figure [-@fig-glm-modfit] shows a plot of the data points along with the three fitted lines. It was obtained using the following code in \textsf{R}, where \texttt{ilogit, iprobit} and \texttt{icloglog} are the inverse of the corresponding link functions : 

```{r} 
#| eval: false 

plot(x,y/m,pch=19,xlab="Number of Hours of Operation (x)", ylab="Prob of Developing Fissures") 
j <- seq(0,4800,1) 
lines(j,ilogit(coef(modelb)[1]+coef(modelb)[2]*j),lwd=2) lines(j,iprobit(coef(modelc)[1]+coef(modelc)[2]*j),lty=2,col=2,lwd=2) lines(j,icloglog(coef(modeld)[1]+coef(modeld)[2]*j),lty=3,col=4,lwd=2) legend("topleft",legend=c("Logit","Probit","Complementary log-log"), lty=c(1,2,3),col=c(1,2,4),lwd=rep(2,3)) 
``` 

```{r} 
#| echo: false 


modelb <- glm(cbind(y,m-y) ~ x,family=binomial) 

modelc <- glm(cbind(y,m-y) ~ x,family=binomial(link=probit)) 

modeld <- glm(cbind(y,m-y) ~ x,family=binomial(link=cloglog)) 
ilogit <- function(x) exp(x)/(1+exp(x)) 
iprobit <- function(x) pnorm(x) 
icloglog <- function(x) 1 - exp(-exp(x)) 

```


```{r}
#| echo: false
#| label: fig-glm-modfit
#| fig-cap: "Logistic, Probit and Complementary Log-Log Model Fit"


plot(x,y/m,pch=19,xlab="Number of Hours of Operation (x)",
     ylab="Prob of Developing Fissures")
j <- seq(0,4800,1)
lines(j,ilogit(coef(modelb)[1]+coef(modelb)[2]*j),lwd=2)
lines(j,iprobit(coef(modelc)[1]+coef(modelc)[2]*j),lty=2,col=2,lwd=2)
lines(j,icloglog(coef(modeld)[1]+coef(modeld)[2]*j),lty=3,col=4,lwd=2)
legend("topleft",legend=c("Logit","Probit","Complementary log-log"),
       lty=c(1,2,3),col=c(1,2,4),lwd=rep(2,3))
```


It is easy to observe that the fit is better when the number of hours of operations is lower, it seems that the variance of the observations is increasing with the predictor. This is expected in a generalized linear model framework. The three fitted lines are slightly different. The probit link produces lower estimates in the tails and higher estimates in the middle of the range of the predictors. It seems like this model is less representative of the data than the others. The complementary log-log model (d) predicts higher probabilities of failures in the extremes of the range of the predictors. This seems to fit the data well, and recall that the variance of the predictions where also smaller than other models in this case, which is a desirable property. The line for the model with canonical link is between the two others. It could also be a reasonable model for the data.


:::

